Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-08 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
More editing of references, Madrzykowski thesis

The source fire characterization experiments were conducted in the NIST Large Fire Research Laboratory’s 3~m by 3~m hood oxygen consumption calorimeter. This apparatus follows the methodology of ASTM E 2067-08, Standard Practice for Full-Scale Oxygen Consumption Calorimetry Fire Tests.[37]  The fires were centered under the hood of the calorimeter and positioned in an open steel frame which provided support for the thermocouples and heat flux sensors. A load cell was positioned as a support for the fuels.  All of the data was recorded at intervals of 1~s on a computer based data acquisition system.

@@

The NIST Large Fire Research Laboratory’s 3~m by 3~m oxygen depletion calorimeter has an estimated peak heat release rate capacity of approximately 700~kW.  Details on the operation and uncertainty in measurements associated with NIST’s 6~m by 6~m oxygen consumption calorimeter were documented by Bryant et al. [38] Many of the same instruments are shared between the two calorimeters and calibration burns were conducted during each experimental series.  As a result, the expanded uncertainty of this device has been estimated as 11~\% on the measured heat release rate.

@@

Gas temperatures were measured with exposed-bead, Chromel-Alumel (type K) thermocouples, with a 1.0~mm nominal diameter.  Starting from the exposed bead, the thermocouple wire was sheathed in a 3.2 mm diameter inconel shield, 0.76~m in length.  The standard uncertainty in temperature of the thermocouple wire itself is 2.2~$^\circ$C at 277~$^\circ$C and increases to 9.5~$^\circ$C at 871~$^\circ$C as determined by the wire manufacturer [39]  The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty.[40, 41]  Small diameter thermocouples were used to limit the impact of radiative heating and cooling.  The estimated combined expanded uncertainty for temperature in these experiments is 15~\%.

@@

Water cooled, Schmidt-Boelter total heat flux gauges were used to measure the heat flux. The sensing surface of the heat flux gauges were located 0.95~m horizontally from the center of the burner/fuel and 0.50~m above the top of the burner/pan lip/fuel surface.  The manufacturer reports a 3~\% calibration expanded uncertainty for these devices 20.[42] Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically 8~\%.[43]  The gauges were mounted to uprights of the steel instrumentation support frame on opposite sides of the fire.

@@

The fires were centered under the hood of the calorimeter on a non-combustible platform measuring 0.91~m x 1.2~m x 12~mm thick supported by the load cell.  The load cell had a resolution of 1.0~g.[44] The expanded uncertainty is estimated to be 5~\%.  The load cell was checked between experiments with calibrated weights during the zero and span checks of the calorimeter.

@@

On each fuel’s averaged heat release rate curve, a bar representing the 95~\% confidence level ($2\sigma$), based on Type A evaluation of the uncertainty is provided.[45]  A Type A evaluation of the uncertainty is any statistical analysis of the data from a series of observations.  In this case the evaluation assumed a normal distribution of the data such that doubling the calculated standard deviation (2σ) would yield a confidence level of approximately 95~\%.

@@

As noted by Beyler in his review of fire plume correlations, video tape analysis seems to provide the best estimates of flame height. [46]  Each experiment was recorded with a digital video camera at a refresh rate frequency of 29.97~fps which is the National Television Standards Committee (NTSC) standard. The resulting image in each frame was 720 pixels wide by 480 pixels high. Segments of the video which corresponded to the peak heat release of each experiment were captured with a video editing program.

@@

Each frame from the video clip was saved as a full color Tagged Image File Format (TIFF) file and the pixel aspect ratio was changed from 0.9 to 1, so that each pixel had the same dimension in both the horizontal and vertical directions.  Once these images were rendered by the video editing program, each image was scaled and cropped to provide input for Streams, a suite of image processing programs developed by Nokes for analyzing fluid flow visualization experiments[26]. The images were scaled so that one pixel equals 5 mm in height and width.

@@

From each of the peak heat release rate video clips, the edited TIFF files were loaded into Streams.  For the natural gas and gasoline fueled fires, approximately 900 images were used in each image sequence.  The polyurethane foam experiments had approximately 450 images in each sequence for each experiment.  Once the image sequence was created, a series of filters and amplifiers are applied to the images to eliminate all colors but red, based on a method developed by Goble.[47]  Once all of the images were converted to a pure red ``flame'' which represented the area of visible fire, on a black background, the images were processed within Streams to develop an intensity field based on all of the images being ``overlaid'' on each other.  The end result is a contour plot of the intermittency of the flame ranging from zero to one.

@@

Another means of quantifying field data with Streams is the use of contour plots.  Examples of the time averaged contour plots are shown in Figures 20 through 22 for a natural gas, gasoline, and polyurethane fueled experiments respectively.  Each contour line represents the percentage of time that the flame was at a given height.  Figure 20 provides an example of a natural gas experiment.  The region under examination, as shown as a black bounding rectangle, is 1400~mm high and 800 mm in width.  This example was averaged over approximately 29~s or approximately 870 images.  The number above each contour line in the plot range from 0 to 1.  The area bounded by the contour with the value 1 represents the area that the flame occupied 100~\% of the time period analyzed or the persistent flame region as defined by McCaffery. [48] The remaining contours are in the intermittent flame region, with the 0.5 contour representing the mean flame height. Outside of the final contour, value 0.0, no visible flame was recorded; hence this contour marks the definitive start of the buoyant plume region.

>>>>

The source fire characterization experiments were conducted in the NIST Large Fire Research Laboratory’s 3~m by 3~m hood oxygen consumption calorimeter. This apparatus follows the methodology of ASTM~E 2067-08, {\em Standard Practice for Full-Scale Oxygen Consumption Calorimetry Fire Tests}~\cite{ASTM_E2067}. The fires were centered under the hood of the calorimeter and positioned in an open steel frame which provided support for the thermocouples and heat flux sensors. A load cell was positioned as a support for the fuels.  All of the data was recorded at intervals of 1~s on a computer based data acquisition system.

@@

The NIST Large Fire Research Laboratory’s 3~m by 3~m oxygen depletion calorimeter has an estimated peak heat release rate capacity of approximately 700~kW.  Details on the operation and uncertainty in measurements associated with NIST’s 6~m by 6~m oxygen consumption calorimeter were documented by Bryant et al.~\cite{Bryant:2004}. Many of the same instruments are shared between the two calorimeters and calibration burns were conducted during each experimental series.  As a result, the expanded uncertainty of this device has been estimated as 11~\% on the measured heat release rate.

@@

Gas temperatures were measured with exposed-bead, Chromel-Alumel (type K) thermocouples, with a 1.0~mm nominal diameter.  Starting from the exposed bead, the thermocouple wire was sheathed in a 3.2~mm diameter inconel shield, 0.76~m in length.  The standard uncertainty in temperature of the thermocouple wire itself is 2.2~$^\circ$C at 277~$^\circ$C and increases to 9.5~$^\circ$C at 871~$^\circ$C as determined by the wire manufacturer~\cite{Omega}.  The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty~\cite{Blevins:1999,Pitts:2001}. Small diameter thermocouples were used to limit the impact of radiative heating and cooling.  The estimated combined expanded uncertainty for temperature in these experiments is 15~\%.

@@

Water cooled, Schmidt-Boelter total heat flux gauges were used to measure the heat flux. The sensing surface of the heat flux gauges were located 0.95~m horizontally from the center of the burner/fuel and 0.50~m above the top of the burner/pan lip/fuel surface.  The manufacturer reports a 3~\% calibration expanded uncertainty for these devices~\cite{Medtherm}. Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically 8~\%~\cite{Pitts:2006}.  The gauges were mounted to uprights of the steel instrumentation support frame on opposite sides of the fire.

@@

The fires were centered under the hood of the calorimeter on a non-combustible platform measuring 0.91~m x 1.2~m x 12~mm thick supported by the load cell.  The load cell had a resolution of 1.0~g~\cite{Mettler}. The expanded uncertainty is estimated to be 5~\%.  The load cell was checked between experiments with calibrated weights during the zero and span checks of the calorimeter.

@@

On each fuel’s averaged heat release rate curve, a bar representing the 95~\% confidence level ($2\sigma$), based on Type A evaluation of the uncertainty is provided~\cite{Taylor:1994}.  A Type A evaluation of the uncertainty is any statistical analysis of the data from a series of observations.  In this case the evaluation assumed a normal distribution of the data such that doubling the calculated standard deviation (2σ) would yield a confidence level of approximately 95~\%.

@@

As noted by Beyler in his review of fire plume correlations, video tape analysis seems to provide the best estimates of flame height~\cite{Beyler:1986}. Each experiment was recorded with a digital video camera at a refresh rate frequency of 29.97~fps which is the National Television Standards Committee (NTSC) standard. The resulting image in each frame was 720 pixels wide by 480 pixels high. Segments of the video which corresponded to the peak heat release of each experiment were captured with a video editing program.

@@

Each frame from the video clip was saved as a full color Tagged Image File Format (TIFF) file and the pixel aspect ratio was changed from 0.9 to 1, so that each pixel had the same dimension in both the horizontal and vertical directions.  Once these images were rendered by the video editing program, each image was scaled and cropped to provide input for Streams, a suite of image processing programs developed by Nokes for analyzing fluid flow visualization experiments~\cite{Nokes:2011}. The images were scaled so that one pixel equals 5~mm in height and width.

@@

From each of the peak heat release rate video clips, the edited TIFF files were loaded into Streams.  For the natural gas and gasoline fueled fires, approximately 900 images were used in each image sequence.  The polyurethane foam experiments had approximately 450 images in each sequence for each experiment.  Once the image sequence was created, a series of filters and amplifiers are applied to the images to eliminate all colors but red, based on a method developed by Goble~\cite{Goble:2007}.  Once all of the images were converted to a pure red ``flame'' which represented the area of visible fire, on a black background, the images were processed within Streams to develop an intensity field based on all of the images being ``overlaid'' on each other.  The end result is a contour plot of the intermittency of the flame ranging from zero to one.

@@

Another means of quantifying field data with Streams is the use of contour plots.  Examples of the time averaged contour plots are shown in Figures 20 through 22 for a natural gas, gasoline, and polyurethane fueled experiments respectively.  Each contour line represents the percentage of time that the flame was at a given height.  Figure 20 provides an example of a natural gas experiment.  The region under examination, as shown as a black bounding rectangle, is 1400~mm high and 800 mm in width.  This example was averaged over approximately 29~s or approximately 870 images.  The number above each contour line in the plot range from 0 to 1.  The area bounded by the contour with the value 1 represents the area that the flame occupied 100~\% of the time period analyzed or the persistent flame region as defined by McCaffrey~\cite{McCaffrey:1979}. The remaining contours are in the intermittent flame region, with the 0.5 contour representing the mean flame height. Outside of the final contour, value 0.0, no visible flame was recorded; hence this contour marks the definitive start of the buoyant plume region.
